---
title: "Life Policy Generation: 10 - Create Life Policy Data"
author: "Mick Cooney"
date: "11 January 2018"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: yes
    theme: cerulean
  pdf_document: default
---


```{r knit_opts, include = FALSE}
knitr::opts_chunk$set(tidy  = FALSE
                     ,cache = FALSE
                     ,message = FALSE
                     ,warning = FALSE
                     ,fig.height =  8
                     ,fig.width  = 11)

library(tidyverse)
library(data.table)
library(dtplyr)
library(gtools)

library(feather)

library(lifecontingencies)

options(width = 80L
       ,warn  = 1
        )

set.seed(421)

source("custom_functions.R")
```

# Load Input Data

This worksheet enables the generation of a fake book of policies for a life
insurance company in Ireland

## Set up input parameters

```{r initialise_parameters, echo=TRUE}
# N is the policy count you want to generate
N <- 1500000


# data_startdate is the first date at which we start creating policies
# data_snapshotdate is the date we take the policy 'snapshot'
data_startdate    <- as.Date('1990-01-01')
data_snapshotdate <- as.Date('2015-12-31')

yearchng_dt <- data.table(year = c(1990, 1994, 1995, 1996, 1997, 1998, 1999, 2008, 2009, 2014)
                         ,chng = c(   1,    1,    1,    1,    1,    1,    1,   -4,   -1,    1)
                          )

month_weight <- c("Jan" =  3, "Feb" = 10, "Mar" = 10, "Apr" = 10, "May" = 10, "Jun" =  5
                 ,"Jul" =  5, "Aug" =  5, "Sep" =  7, "Oct" =  7, "Nov" =  7, "Dec" =  2)

dow_weight <- c("Mon" = 5, "Tue" = 7, "Wed" = 7, "Thu" = 7, "Fri" = 5, "Sat" = 0, "Sun" = 0)

holiday_datestr <- c("0101", "0317", "1225", "1226")


# prem_freq is distributed such that almost all are monthly payments
# though a few are yearly
prem_freq_prop   <- c("12" = 0.9, "4" = 0.02, "1" = 0.08)
prem_freq_cutoff <- as.Date('2005-01-01')

# Select which cluster count we are using for our policy generation
use_cluster_level = 'n6'


# Parameters for the gamma distribution used to generate the prem_ape variable
# Protection policies will later be calculated for protection policies
premape_shape <- 1.1
premape_scale <- 500


# Set the gender proportion for the book and different smoking levels
# In order, the labels are S, Q, and N
male_prop   <- 0.7
smoker_prop <- list("M" = c(0.30, 0.20, 0.50)
                   ,"F" = c(0.25, 0.40, 0.35))

# Set proportions for various protection policies
jointlife_prop       <- 0.3
lifeonly_prop        <- 0.75
large_amount_prop    <- 0.05

mortgage_status_prop <- c("TERM" = 0.4, "MORTDECR" = 0.5, "MORTFULL" = 0.1)

# Add 40% to the premium to account for profit margin and expenses
premium_expense_ratio <- 0.4


# Interest rate for discounting time value of money
interest_rate <- 0.04

# We load some lifetable and morbidity data to price the protection policies
load("data/lifemorb_tables.rda")

act_tables <- list(m_lifetable       = m_lifetable
                  ,f_lifetable       = f_lifetable
                  ,mn_morb_lifetable = mn_morb_lifetable
                  ,fn_morb_lifetable = fn_morb_lifetable)

output_csv      <- sprintf("dt_policy_%08d.csv", N)
output_feather  <- sprintf("dt_policy_%08d.feather", N)
output_json     <- sprintf("lifeins_policybook_%08d.json", N)
output_rds      <- sprintf("lifeins_policybook_%08d.rds", N)

output_intermediate_rds  <- sprintf("lifeins_policybook.rds", N)

print(paste(output_csv))
```

## Load Data from Files


```{r load_datafiles, echo=TRUE}
# Load up the smallarea data that comes with the census GIS data
dt_smallarea <- fread("data/smallarea_gisdb.csv")[, sa_id := geogid]

# Load up the clustering info for the small areas
dt_clustersmallarea <- melt(fread("data/aai_censusclusterids.csv"), id = 'sa_id')[
    ,cluster_id := paste0(variable, "_c", value)][
    ,.(sa_id, cluster_level = variable, cluster_id)]

# Load the cluster proportions in the policy book
dt_clusterpolicyprop <- fread("data/cluster_policy_proportions.csv")

# Load the cluster propensities for the four product types
dt_clusterproductmapping <- fread("data/cluster_product_mapping.csv")

# Load product data
dt_productdata    <- fread("data/product_input_data.csv")
dt_productdurprop <- fread("data/product_duration_data.csv")
```

# Create Policy Data

We need to generate a policyid which is just a unique identifier

```{r generate_policy_ids, echo=TRUE}
policyid_N <- 10^floor(log10(N))
policy_id <- sample((policyid_N*10):(policyid_N*100), N, replace = FALSE)

policy_id <- sprintf("C%09d", policy_id)
```

First randomly select a cluster for the policy and use this to set a small area

```{r create_cluster_data, echo=TRUE}
dt_smallarea <- merge(dt_smallarea
                      ,dt_clustersmallarea[cluster_level %in% use_cluster_level]
                      ,by = 'sa_id')

dt_policyprop  <- dt_clusterpolicyprop[cluster_level == use_cluster_level]
cluster_count  <- dt_policyprop[, .N]

cluster_idx <- sample(1:cluster_count, N, prob = dt_policyprop$proportion, replace = TRUE)
cluster_id  <- dt_policyprop[cluster_idx]$cluster_id


dt_clusterprod <- melt(dt_clusterproductmapping
                       ,id = c("cluster_level", "cluster_id", "cluster_name"))

dt_policy <- rbindlist(lapply(unique(cluster_id), sample_cluster_data, dt_smallarea, dt_clusterprod, cluster_id))
dt_policy[, policy_id := policy_id]
```

```{r generate_product_data, echo=TRUE}
### For each prod_type we create some data and then bind the data.tables together
dt_product <- rbindlist(lapply(dt_policy[, sort(unique(prod_type))]
                              ,sample_product_data
                              ,dt_productdata
                              ,premape_shape = premape_shape
                              ,premape_scale = premape_scale
                              ))

dt_product[, policy_id := dt_policy[prod_type == .BY$prod_type]$policy_id, by = prod_type]
dt_policy[,  prod_type := NULL]

dt_policy <- merge(dt_policy, dt_product, by = 'policy_id')
```

We now want to build a model for how policies are created over time.

```{r generate_policy_startdates, echo=TRUE}
allyears_dt <- data.table(year = year(data_startdate):year(data_snapshotdate))

year_dt <- allyears_dt %>%
    left_join(yearchng_dt, by = 'year') %>%
    mutate(chng   = ifelse(is.na(chng), 0, chng)
          ,weight = cumsum(chng)
           )

month_dt <- data.table(month  = names(month_weight)
                      ,weight = month_weight
                       )

dow_dt <- data.table(dow    = names(dow_weight)
                    ,weight = dow_weight
                     )

startdates <- generate_random_startdates(nrow(dt_policy)
                                        ,data_startdate
                                        ,data_snapshotdate
                                        ,holiday_datestr
                                        ,year_dt
                                        ,month_dt
                                        ,dow_dt)

dt_policy[, policy_startdate := startdates]


### Create some datetime values

dt_policy[, dob_life1 := policy_startdate - (365 * policy_startage)]
dt_policy[, dob_life1 := dob_life1 - sample(1:365, N, replace = TRUE) + 1]
```


Having created the start date for the policy and some other derived
date variables from that, we determine the premium type for the policy
and also ensure that after a certain date, quarterly premium policies
are no longer generated.


```{r generate_gender_smoker_data, echo=TRUE}
### Select the prem_freq for RP policies
dt_policy[prem_type == "RP", prem_freq := sample(names(prem_freq_prop)
                                                ,.N
                                                ,replace = TRUE
                                                ,prob    = prem_freq_prop)]

dt_policy[prem_type == 'RP' &
          prem_freq == '4'  &
          policy_startdate >= prem_freq_cutoff, prem_freq := '12']

### Create gender and gender-related variables
dt_policy[, gender_life1 := sample(c("M", "F")
                                  ,N
                                  ,replace = TRUE
                                  ,prob = c(male_prop, 1 - male_prop))]

dt_policy[gender_life1 == 'M'
         ,smoker_life1 := sample(c("S", "Q", "N")
                                ,.N
                                ,replace = TRUE
                                ,prob = smoker_prop[["M"]])]

dt_policy[gender_life1 == 'F'
         ,smoker_life1 := sample(c("S", "Q", "N")
                                ,.N
                                ,replace = TRUE
                                ,prob = smoker_prop[["F"]])]
```

Estimate the policy duration

```{r determine_policy_duration, echo=TRUE}
dt_duration <- dt_policy[, {
    use_dt <- dt_productdurprop[prod_type == .BY$prod_type]

    if(nrow(use_dt) > 0) {
        policy_duration <- sample(use_dt$duration,.N,replace = TRUE,prob = use_dt$prop)
    } else {
        policy_duration <- rep(as.integer(NA), .N)
    }

    .(policy_id = policy_id, policy_duration = policy_duration)
}, by = prod_type]

dt_policy <- merge(dt_policy, dt_duration[, .(policy_id, policy_duration)], by = 'policy_id')
dt_policy[policy_duration == -1, policy_duration := NA]

# Add some sanity checks for policy holder age - No-one should be older than 65,
# and no policies cover beyond the age of 75
dt_policy[prod_type == 'protection' & policy_startage > 65, policy_duration := 5]
dt_policy[prod_type == 'protection' & (policy_startage + policy_duration > 75)
         ,policy_duration := as.integer(75 - policy_startage)]

# Policies starting before 2002-01-01 do not have a policy_duration of longer than 20 years
dt_policy[prod_type        == 'protection' &
          policy_startdate < as.Date('2002-01-01') &
          policy_duration  > 20
         ,policy_duration := 20]


# If a policy has no duration we set the end date to be when the holder turns 120 years old
# Otherwise we set the end date to be the date when the policy expires
dt_policy[is.na(policy_duration)
         ,policy_enddate := lubridate::`%m+%`(dob_life1, lubridate::years(120))]

# The bizarre syntax for %m+% is because I did not want to load lubridate as it overwrites
# some data.table stuff. Stupid R namespace management
dt_policy[!is.na(policy_duration)
         ,policy_enddate := lubridate::`%m+%`(policy_startdate
                                             ,lubridate::years(policy_duration))]

ggplot(dt_policy %>% filter(prod_type == 'protection', policy_startage >= 55)) +
    geom_point(aes(x = policy_startage, y = policy_duration))


### Add some extra data for mortgage-linked protection policies and afew other pieces of information
dt_policy[prod_type == 'protection'
         ,mortgage_status := sample(names(mortgage_status_prop)
                                   ,.N
                                   ,replace = TRUE
                                   ,prob = mortgage_status_prop)]

dt_policy[prod_type == 'protection', isjointlife := rbinom(.N, 1, jointlife_prop) == 1]
dt_policy[prod_type == 'protection', islifeonly  := rbinom(.N, 1, lifeonly_prop)  == 1]


### Add a mortality rating to protection policies
dt_policy[prod_type == 'protection'
         ,mort_rating := calculate_mort_rating(.N
                                              ,dt_clusterpolicyprop[cluster_id == .BY$cluster_id, mort_shape]
                                              ,dt_clusterpolicyprop[cluster_id == .BY$cluster_id, mort_scale])
         ,by = cluster_id]

dt_policy[prod_type == 'protection' & smoker_life1 == "S", mort_rating := mort_rating + 100]
dt_policy[prod_type == 'protection' & smoker_life1 == "Q", mort_rating := mort_rating + 50]

t(head(dt_policy, 3))
```

Calculate the premium for the protection policies

```{r calculate_policy_premia, echo=TRUE}
dt_policy[, policy_startage := as.numeric(floor((policy_startdate - dob_life1) / 365))]

dt_policy[prod_type == 'protection'
         ,premium_factor := calculate_protection_premium(policy_startage
                                                        ,policy_duration
                                                        ,isjointlife
                                                        ,islifeonly
                                                        ,acttables     = act_tables
                                                        ,interest_rate = interest_rate)
         ,by = policy_id]

dt_policy[prod_type == 'protection'
         ,sum_assured := 100000 + 50000 * sample(0:8, .N, replace = TRUE)]

dt_policy[prod_type == 'protection'
         ,sum_assured := ifelse(runif(.N) < large_amount_prop, sum_assured * 10, sum_assured)]

dt_policy[prod_type == 'protection'
         ,prem_risk := (sum_assured * premium_factor * mort_rating / 100)]

dt_policy[prod_type == 'protection' & mortgage_status == 'MORTDECR'
         ,prem_risk := prem_risk * 0.7]
```

```{r add_expenses_profit_margin, echo=TRUE}
### Once we have calculated a premium, we need to add a margin for
### expenses and for profit margin etc.
### We have set this as an additional 30% to the premium

dt_policy[prod_type == 'protection', prem_ape  := round(prem_risk * (1 + premium_expense_ratio), 2)]

dim(dt_policy[prod_type == 'protection' & is.na(policy_duration)])
```


```{r generate_geographical_data, echo=TRUE}
### Add some geographical information
dt_policy <- merge(dt_policy, dt_smallarea[, .(sa_id, countyname, edname, nuts3name)]
                   ,by = 'sa_id',all_x = TRUE)

dt_output <- dt_policy[, .(policy_id,countyname,edname,nuts3name,sa_id,cluster_id,prod_type
                           ,prem_type,prem_freq,prem_ape,prem_risk,policy_startdate,policy_enddate
                           ,policy_duration,mort_rating,sum_assured
                           ,dob_life1,gender_life1,smoker_life1,isjointlife,islifeonly
                           ,mortgage_status)]

setkey(dt_output, policy_id)

dt_output[, quantile(prem_ape, 0.50, na.rm = TRUE), by = prod_type]

ggplot(dt_output %>% filter(prod_type == 'protection')) +
    geom_density(aes(x = prem_ape))
```


# Output Data

Write the CSV data out to files

```{r output_data, echo=TRUE}
write_csv    (dt_output, path = paste0('output/', output_csv))
write_feather(dt_output, path = paste0('output/', output_feather))

# Convert the data to JSON and output it
#write(jsonlite::toJSON(dt_output), file = paste0('output/', output_json))

saveRDS(dt_output, file = paste0('output/',output_rds),              compress = 'xz')
saveRDS(dt_output, file = paste0('output/',output_intermediate_rds), compress = 'xz')


```
